---
title: "FA_Cup_Magic"
author: "Ye Yint Win"& "Jaleta"
date: "2023-05-03"
output: html_document
---

# Is the FA Cup Magic Real?

## Library

Add tidyverse, dplyr and MASS library

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(MASS)
```

## Reading CSV

First, go to this website <https://github.com/jalapic/engsoccerdata>.

Click on data_raw, and then click on england.csv and click on "view raw" at the bottom and download the .csv file. Do the same thing to facup.csv.

read_csv for both fa_cup and england data

Filter Season to \>=1958 and round to (6,s,f) in fa_cup data.

```{r}
fa_cup<-read_csv("https://raw.githubusercontent.com/jalapic/engsoccerdata/master/data-raw/facup.csv")#Getting Fa_Cup CSV
#Filtering the Season because the raw data and English Football Association created a fourth division rather than thrid division(North and South)
fa_cup<- fa_cup %>%
  filter(
    Season>= 1958,
    round %in% c('6','s','f')
  )
fa_cup

```

Then, filter the england.csv to Season to \>=1958 and \<= 2016 to get the same Seasons with facup.csv.

Then, create a new table, new_england, and do distinct with (home, Season, and division) and group by (Season and home). Make a new table, total_goals by grouping by the same thing and use summarise and sum of hgoal (homegoal).

Now, left_join total_goals to the new_england table by the keys Season and home.

```{r}
#Getting All time Official english football match data
#Filtering the data to match the FA Cup Data
england<-read_csv("https://raw.githubusercontent.com/jalapic/engsoccerdata/master/data-raw/england.csv")
england <-england%>%
  filter(
    Season >=1958 & Season <=2016
  )
england


#Getting only Season, home and Division to get leftjoin on Facup data
new_england <- england %>%
  distinct(home, Season, division)%>%
  group_by(Season,home)

total_goals <- england %>%
  group_by(Season,home) %>%
  summarize(total_goals = sum(hgoal))
total_goals

new_england <- left_join(new_england, total_goals, by = c("Season","home"))
new_england
```

## Left-Joins

Create a new table, join_facup_home, to get division and totalgoal number for the home team. Here, leftjoing by home and seasona, group by Season, and mutate division and total_goal to hometier and home_totalgoal so that it will be different from visitor's. Then, select the useful columns.

Do the same for the new table for visitor team, join_facup_vis. But, here, don't select all columns, but only Season, home, visitor, visittier and vis_totalgoal so that the columns will not repeat in the leftjoin again later. (use dplyr:: select instead of just select).

Lastly, create the last new table tier_facup to leftjoin the home table earlier from the visitor table earlier with the keys of Season, home, visitor. Then, select useful columns.

```{r}
#Left-join to FA_Cup to get which division the home team is in 
join_facup_home<- left_join(fa_cup, new_england, by = c("home", "Season"))%>%
  group_by(Season)%>%
  mutate(hometier = division,home_totalgoal = total_goals) %>%
  dplyr::select(Season:round, attendance, hometier,home_totalgoal)
#Left-join to FA_Cup to get which division the visitor team is in 
join_facup_vis<- left_join(fa_cup, new_england, by = c("visitor"= "home", "Season"))%>%
  group_by(Season)%>%
  mutate(visittier = division,vis_totalgoal = total_goals)%>%
  dplyr::select(Season, home,visitor,visittier,vis_totalgoal)

join_facup_home
join_facup_vis


#Left join to the home table from visitor team so that we can have division data for both sides
tier_facup<- left_join(join_facup_home,join_facup_vis, by = c( "Season", "home","visitor"))
tier_facup<- tier_facup%>%
  dplyr::select(Season: hometier, visittier,home_totalgoal,vis_totalgoal)
  
tier_facup


```

## Mutate in Wrangled Data

```{r}

#
tier_facup<- tier_facup %>%
  mutate( 
  match_won = hgoal-vgoal, #Match winner
  match_won = ifelse(match_won>0, "H", ifelse(match_won<0, "V","T")), # Determining Winner or Tie
  tier_diff = as.numeric(hometier)-as.numeric(visittier), #getting Division/Tier Difference
  tier_diff = ifelse(tier_diff<0, tier_diff*(-1), tier_diff),#Making all tier_diff positives
  magic = case_when
         (
    as.numeric(hometier)<as.numeric(visittier) & hgoal>vgoal | 
    as.numeric(hometier)>as.numeric(visittier) & hgoal<vgoal |   
    as.numeric(hometier)>as.numeric(visittier) & hgoal == vgoal | 
    as.numeric(hometier)<as.numeric(visittier) & hgoal==vgoal ~ 1, 
    TRUE ~ 0
  ),
  goal_diff = case_when
         (
    as.numeric(hometier)<as.numeric(visittier) & hgoal>vgoal | 
    as.numeric(hometier)>as.numeric(visittier) & hgoal<vgoal |   
    as.numeric(hometier)>as.numeric(visittier) & hgoal == vgoal | 
    as.numeric(hometier)<as.numeric(visittier) & hgoal==vgoal ~ as.numeric(home_totalgoal)-as.numeric(vis_totalgoal), 
    TRUE ~ NA_real_),
  goal_diff= ifelse(goal_diff<0, goal_diff*(-1),goal_diff),
  real_goal_diff = as.numeric(home_totalgoal)-as.numeric(vis_totalgoal),
  real_goal_diff = ifelse(real_goal_diff<0, real_goal_diff*(-1),real_goal_diff)
  )

tier_facup
```

```{r}

tier_facup<-tier_facup%>% 
  mutate(tier_diff = as.numeric(as.character(tier_diff)))
tier_facup

```

```{r}
mod4 <- glm(magic ~ real_goal_diff + as.factor(tier_diff) +real_goal_diff : as.factor(tier_diff), data = tier_facup, family= "binomial") 
summary(mod4)
mod.4<- coef(mod4)

mean_rgd <- mean(tier_facup$real_goal_diff[!is.na(tier_facup$real_goal_diff)])
mean_td <- mean(as.numeric(tier_facup$tier_diff[!is.na(tier_facup$tier_diff)]))

mean_rgd
mean_td

func.4.rdg <- function(x){plogis(mod.4[1]+mod.4[2]*x+mod.4[3]*mean_td)}
ggplot(tier_facup) +
  geom_jitter( aes(x = real_goal_diff,y=magic), height = 0.05,alpha = 1) +
  geom_function(fun = func.4.rdg, color = "red")

func.4.td1 <- function(x){plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[3]*x+mod.4[6]*x)}
func.4.td2 <- function(x){plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[4]*x+mod.4[7]*x)}
func.4.td3 <- function(x){plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[5]*x+mod.4[8]*x)}
func.4.td0 <- function(x){plogis(mod.4[1]+mod.4[2]*mean_rgd+x*0)}


# Plot the functions
ggplot(tier_facup) +
  geom_jitter(aes(x = tier_diff, y = magic), height = 0.05, alpha = 1) +
  geom_function(fun = func.4.td1, color = "red") +
  geom_function(fun = func.4.td2, color = "blue") +
  geom_function(fun = func.4.td3, color = "darkgreen") +
  geom_function(fun = func.4.td0, color = "black")+
  labs(title = "Probabilty of Lower-division team winning using goal difference in their normal season and Division")


1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[3])
1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[3]+ mod.4[6])
1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[4]+ mod.4[7])
1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[5]+ mod.4[8])

exp(1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[3]))
exp(1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[3]+ mod.4[6]))
exp(1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[4]+ mod.4[7]))
exp(1-plogis(mod.4[1]+mod.4[2]*mean_rgd+mod.4[5]+ mod.4[8]))


```

```{r}
mod5 <- glm(magic ~ real_goal_diff +attendance+ as.factor(tier_diff), data = tier_facup, family= "binomial") 
summary(mod5)
mod.5<- coef(mod5)


mean_att<- mean(as.numeric(tier_facup$attendance[!is.na(tier_facup$attendance)]))
mean_att 


func.3.rdg <- function(x){plogis(mod.3[1]+mod.3[2]*x+mod.3[3]*mean_td)}
ggplot(tier_facup) +
  geom_jitter( aes(x = real_goal_diff,y=magic), height = 0.05,alpha = 1) +
  geom_function(fun = func.3.rdg, color = "red")

func.5.td1 <- function(x){plogis(mod.5[1]+mod.5[2]*mean_rgd+mod.5[3]* mean_att+ mod.5[4]*x)}
func.5.td2 <- function(x){plogis(mod.5[1]+mod.5[2]*mean_rgd+mod.5[3]* mean_att+ mod.5[5]*x)}
func.5.td0 <- function(x){plogis(mod.5[1]+mod.5[2]*mean_rgd+mod.5[3]* mean_att+x*0)}

ggplot(tier_facup) +
  geom_jitter( aes(x = tier_diff,y=magic), height = 0.05,alpha = 1) +
  geom_function(fun = func.5.td0, color = "red")+
  geom_function(fun = func.5.td1, color = "blue")+
  geom_function(fun = func.5.td2, color = "darkgreen")

exp(1-plogis(mod.5[1]+mod.5[2]*mean_rgd+mod.5[3]* mean_att))
exp(1-plogis(mod.5[1]+mod.5[2]*mean_rgd+mod.5[3]* mean_att+mod.5[4]))
exp(1-plogis(mod.5[1]+mod.5[2]*mean_rgd+mod.5[3]* mean_att+ mod.5[5]))


```

```{r}
mod4$aic
mod5$aic

anova(mod4, test="Chisq")
anova(mod5, test="Chisq")
```
